### 2024-12-04 draft.tex
Done informal proof of CLT

\subsection{Central Limit Theorem}
Informal proof of Lyapounov's theorem for a single sequence by the method of Lindeberg \cite{chung2000course}.
\begin{itemize}
    \item \textbf{Notation and Statement}
    \begin{enumerate}
    \item \( \{X_j\}_{j=1}^n \) be a sequence of independent random variables with:
    \begin{itemize}
        \item[] \( \mathbb{E}[X_j] = 0 \), 
        \item[] \( \text{Var}(X_j) = \sigma_j^2 \) (finite variances, \( \sigma_j^2 < \infty \)),
        \item[] \( \mathbb{E}[|X_j|^3] = \gamma_j \) (finite third moments, \( \gamma_j < \infty \)).
    \end{itemize}

@@

        \item Statement:
    If 
    \[
   \frac{\Gamma_n}{s_n^3} \to 0 \quad \text{as } n \to \infty.
   \]
   then \( \frac{S_n}{s_n} \xrightarrow{d} \mathcal{N}(0,1) \),
   where \( \mathcal{N}(0,1) \) denotes the standard normal distribution.

@@

Now, construct the sequence:
\[
Z_j = Y_1 + \dots + Y_{j-1} + X_j + \dots + X_n, \quad 1 \leq j \leq n.
\]
Thus:
\begin{itemize}
    \item[] \( Z_1 = X_2 + X_3 + \dots + X_n \),
    \item[] \( Z_2 = Y_1 + X_3 + \dots + X_n \),
    \item[] \( Z_n = Y_1 + Y_2 + \dots + Y_{n-1} \).
\end{itemize}

@@

To show that \( \frac{S_n}{s_n} \xrightarrow{d} \mathcal{N}(0, 1) \), we compare the expectations of \( \frac{S_n}{s_n} \) with those of \( \frac{Z_j}{s_n} \).

***** cont. *****

>>>>

\newpage
\subsection{Central Limit Theorem}
Informal proof of Lyapounov's theorem for a single sequence by the method of Lindeberg \cite{chung2000course}.
\begin{itemize}
    \item \textbf{Notation and Statement}
    \begin{enumerate}
    \item \( \{X_j\}_{j=1}^n \) be a sequence of independent random variables with:
    \begin{itemize}
        \item[] \( \mathbb{E}[X_j] = 0 \), 
        \item[] \( \text{Var}(X_j) = \sigma_j^2 \) (finite variances, \( \sigma_j^2 < \infty \)),
        \item[] \( \mathbb{E}[|X_j|^3] = \gamma_j \) (finite third moments, \( \gamma_j < \infty \)).
    \end{itemize}

@@

        \item Statement:
    If 
    \[
   \frac{\Gamma_n}{s_n^3} \to 0 \quad \text{as } n \to \infty
   \]
   then 
   \[
   \frac{S_n}{s_n} \xrightarrow{d} \Phi
   \]
   where \( \Phi\) denotes the standard normal distribution \( \mathcal{N}(0,1) \).

@@

Since  \( Y_j \sim \mathcal{N}(0, \sigma_j^2) \), the normalized random variable:
\[
\frac{Y_j}{s_n} \sim \mathcal{N}\left(0, \frac{\sigma_j^2}{s_n^2}\right)
\]

Adding these normalized variables gives:
\[
\frac{1}{s_n} \sum_{j=1}^n Y_j \sim \mathcal{N}\left(0, \sum_{j=1}^n \frac{\sigma_j^2}{s_n^2}\right).
\]

Since \( s_n^2 = \sum_{j=1}^n \sigma_j^2 \), this simplifies to:
\[
\frac{1}{s_n} \sum_{j=1}^n Y_j \sim \mathcal{N}(0, 1).
\]

Now, construct the sequence \( Z\):
\[
Z_j = Y_1 + \dots + Y_{j-1} + X_j + \dots + X_n, \quad 1 \leq j \leq n
\]
Thus:
\begin{itemize}
    \item[] \( Z_1 = X_2 + X_3 + \dots + X_n \),
    \item[] \( Z_2 = Y_1 + X_3 + \dots + X_n \),
    \item[] \( Z_n = Y_1 + Y_2 + \dots + Y_{n-1} \)
\end{itemize}

@@

To show that \( \frac{S_n}{s_n} \xrightarrow{d} \mathcal{N}(0, 1) \), we compare their expectations.

\textbf{Using Test Funtions}

Consider a test function \( f \) from \( C^3_B \), the class of bounded continuous functions with three bounded derivatives.

Consider:
\[
\mathbb{E}\left[f\left(\frac{S_n}{s_n}\right)\right] \quad \text{and} \quad \mathbb{E}\left[f\left(\mathcal{N}\right)\right],
\]

By introducing the replacement sequence \( Z_j \), we rewrite:

\[
\mathbb{E}\left[f\left(\frac{S_n}{s_n}\right)\right] - \mathbb{E}\left[f\left(\frac{Z_n}{s_n}\right)\right].
\]

By telescoping:
\[
\mathbb{E}\left[f\left(\frac{S_n}{s_n}\right)\right] - \mathbb{E}\left[f\left(\frac{Z_n}{s_n}\right)\right] = \sum_{j=1}^n \left[\mathbb{E}\left[f\left(\frac{X_j + Z_j}{s_n}\right)\right] - \mathbb{E}\left[f\left(\frac{Y_j + Z_{j+1}}{s_n}\right)\right]\right]
\]

\item \textbf{Taylor Expansion of \( f \)}

By Taylor's theorem for \( f \in C^3_B \):
\[
\left|f(x + y) - \left[f(x) + f'(x)y + \frac{f''(x)}{2}y^2\right]\right| \leq \frac{M}{6}|y|^3,
\]
where \( M = \sup_{x \in \mathbb{R}} |f^{(3)}(x)| \) is the maximum bound on the third derivative of \( f \).

Apply this to \( f(\xi + \eta) \) and take Expectations, we write:
\[
\left| \mathbb{E}[f(\xi + \eta)] - \mathbb{E}[f(\xi)] - \mathbb{E}[f'(\xi)]\mathbb{E}[\eta] - \frac{1}{2}\mathbb{E}[f''(\xi)]\mathbb{E}[\eta^2] \right| \leq \frac{M}{6} \mathbb{E}[|\eta|^3],
\]
where \( \xi \) and \( \eta \) are independent random variables such that \(\mathbb{E}[|\eta|^3] < \infty \).

Then, let \( \zeta \) be another independent random variable, having the same mean and variance as \( \eta \). Now, consider:
     \[
     \mathbb{E}[f(\xi + \eta)] - \mathbb{E}[f(\xi + \zeta)].
     \]

Using Taylor's theorem again, this difference can be bounded as:
     \[
     \left| \mathbb{E}[f(\xi + \eta)] - \mathbb{E}[f(\xi + \zeta)] \right| \leq \frac{M}{6} (\mathbb{E}[|\eta|^3] + \mathbb{E}[|\zeta|^3]),
     \]
     where the sum \( \mathbb{E}[|\eta|^3] + \mathbb{E}[|\zeta|^3] \) accounts for the third absolute moments of both random variables.

Then, substitute with:
\begin{itemize}
    \item \( \xi = Z_j / s_n \),
    \item \( \eta = X_j / s_n \),
    \item \( \zeta = Y_j / s_n \).
\end{itemize}
We have:
   \begin{equation} \label{eq:1}
       \left| \mathbb{E}\left[f\left(\frac{Z_j}{s_n} + \frac{X_j}{s_n}\right)\right] - \mathbb{E}\left[f\left(\frac{Z_j}{s_n} + \frac{Y_j}{s_n}\right)\right] \right| \leq \frac{M}{6} \left( \mathbb{E}\left[\left| \frac{X_j}{s_n} \right|^3\right] + \mathbb{E}\left[\left| \frac{Y_j}{s_n} \right|^3\right] \right)
   \end{equation}

\item \textbf{Error Bound Using Lyapunov's Condition}

From the definition of \( X_j / s_n \) and \( Y_j / s_n \), we know:
     \[
     \mathbb{E}\left[\left|\frac{X_j}{s_n}\right|^3\right] = \frac{\mathbb{E}[|X_j|^3]}{s_n^3}, \quad \mathbb{E}\left[\left|\frac{Y_j}{s_n}\right|^3\right] = \frac{\mathbb{E}[|Y_j|^3]}{s_n^3}.
     \]

For \( Y_j \sim \mathcal{N}(0, \sigma_j^2) \), the third absolute moment is:
     \[
     \mathbb{E}[|Y_j|^3] = c \sigma_j^3,
     \]
     where \( c = \sqrt{8 / \pi} \) (a constant depending on the normal distribution).

Since \( \mathbb{E}[|X_j|^3] = \gamma_j \) and \(\mathbb{E}[|Y_j|^3] = c \sigma_j^3\), we wrire:
     \[
   \mathbb{E}\left[\left|\frac{X_j}{s_n}\right|^3\right] + \mathbb{E}\left[\left|\frac{Y_j}{s_n}\right|^3\right] = \frac{\gamma_j}{s_n^3} + \frac{c \sigma_j^3}{s_n^3}
   \]

By summing over \(j = 1, \dots , n\), \eqref{eq:1} becomes:
\[
\sum_{j=1}^n \left| \mathbb{E}\left[f\left(\frac{Z_j}{s_n} + \frac{X_j}{s_n}\right)\right] - \mathbb{E}\left[f\left(\frac{Z_j}{s_n} + \frac{Y_j}{s_n}\right)\right] \right| \leq \frac{M}{6} \sum_{j=1}^n \left( \frac{\gamma_j}{s_n^3} + \frac{c \sigma_j^3}{s_n^3} \right)
\]

Lyapunov's inequality states that:
\[
\sigma_j^3 \leq \gamma_j.
\]
Thus:
\[
\frac{\gamma_j}{s_n^3} + \frac{c \cdot \sigma_j^3}{s_n^3} \leq \frac{\gamma_j}{s_n^3} + \frac{c \cdot \gamma_j}{s_n^3} = \frac{(1 + c) \cdot \gamma_j}{s_n^3}.
\]
Simplify the Sum, we have:

\begin{equation} \label{eq:2}
\frac{M}{6} \sum_{j=1}^n \left( \frac{\gamma_j}{s_n^3} + \frac{c \cdot \sigma_j^3}{s_n^3} \right) \leq \frac{M}{6} \cdot \frac{1}{s_n^3} \sum_{j=1}^n (1 + c) \gamma_j
\end{equation}

Let \( K = \frac{M}{6}(1 + c) \), and the total third absolute moment of \( X_j \):
\[
\Gamma_n = \sum_{j=1}^n \gamma_j ( \space as \space define \space above)
\]
We substitute these into the inequality \eqref{eq:2}:
\begin{equation} \label{eq:3}
\frac{M}{6} \sum_{j=1}^n \left( \frac{\gamma_j}{s_n^3} + \frac{c \cdot \sigma_j^3}{s_n^3} \right) \leq K \cdot \frac{\Gamma_n}{s_n^3}
\end{equation}

By the definition of Big-O Notation, we have 
\begin{displayquote}
For a function \( f(n) \), we say that \( f(n) = O(g(n)) \) if there exist positive constants \( C > 0 \) and \( n_0 \geq 1 \) such that:
\[
|f(n)| \leq C |g(n)|, \quad \text{for all } n \geq n_0
\]
\end{displayquote}

Here:
\begin{itemize}
    \item \( f(n) = \frac{\Gamma_n}{s_n^3} \),
    \item \( g(n) = \frac{\Gamma_n}{s_n^3} \)
\end{itemize}

Clearly:
\[
|f(n)| = \left| \frac{\Gamma_n}{s_n^3} \right| \quad \text{and} \quad |g(n)| = \left| \frac{\Gamma_n}{s_n^3} \right|
\]

Since \( f(n) \) and \( g(n) \) are identical, we have:
\[
|f(n)| \leq 1 \cdot |g(n)| \quad \text{for all } n \geq 1
\]

Here, the constant \( C = 1 \). Then, the inequality \( |f(n)| \leq C |g(n)| \) holds for all \( n \geq 1 \), so we can choose \( n_0 = 1 \).

Since we have found constants \( C = 1 \) and \( n_0 = 1 \) such that:
\[
|f(n)| \leq C |g(n)| \quad \text{for all } n \geq n_0,
\]
it follows by the definition of Big-O notation that:
\begin{equation} \label{eq:4}
\frac{\Gamma_n}{s_n^3} = O\left(\frac{\Gamma_n}{s_n^3}\right)
\end{equation}

The Big O Notation property called \textbf{Multiplication by a constant} states that:
Let \( k \) be a nonzero constant. Then:
\[
O(|k| \cdot g) = O(g)
\]
Apply this to \eqref{eq:4}, we have:
   \[
   K \cdot \frac{\Gamma_n}{s_n^3} = O\left(\frac{\Gamma_n}{s_n^3}\right)
   \]
Then \eqref{eq:3} becomes:
\[
\frac{M}{6} \sum_{j=1}^n \left( \frac{\gamma_j}{s_n^3} + \frac{c \cdot \sigma_j^3}{s_n^3} \right) \leq O\left(\frac{\Gamma_n}{s_n^3}\right)
\]

We have thus obtained the following estimate:
\[
\forall f \in C^3: \left| \space  \mathbb{E}\left[f\left(\frac{S_n}{s_n}\right)\right] - \mathbb{E}\left[f\left(N\right)\right] \right| \leq O\left(\frac{\Gamma_n}{s_n^3}\right) 
\]

\item \textbf{Convergence to Normal}

By Lyapunov's condition:
\[
\frac{\Gamma_n}{s_n^3} \to 0 \quad \text{as } n \to \infty
\]
This directly implies that:
\[
\left| \mathbb{E}\left[f\left(\frac{S_n}{s_n}\right)\right] - \mathbb{E}\left[f(N)\right] \right| \to 0.
\]

Since \( \mathbb{E}[f(S_n / s_n)] \to \mathbb{E}[f(N)] \) for all bounded and continuous test functions \( f \), by the general criterion for vague convergence, we conclude that:
\[
\frac{S_n}{s_n} \xrightarrow{d} \mathcal{N}(0, 1)
\]
